Windowed least-squares model reduction for dynamical systems

### 2019-10-22 section4/spacetime_numerical.tex
through fixes

\subsection{\spaceTimeAcronym\ trial subspaces: direct solution approach}
\KTC{Fix this section:
\begin{enumerate} 
\item remove $\theta^\star$ \EP{I think we still need this for Eq. 38}
 \end{enumerate}
}
The direct solution technique seeks to minimize the fully discrete objective
function associated with the FOM O$\Delta$E. We consider linear
multistep methods and leverage the discretization introduced in
Section~\ref{sec:direct}. For notational simplicity, we define an
index-mapping function that is equivalent to the mapping function
$\indexMapper$, but outputs only the first argument: 
$$\indexMappern: (n,i) \mapsto 
\begin{cases}
n & n = 1, \; i = 0, \\
n & n \ge 1, \; i > 0, \\
\indexMappern(n-1,\nstepsArg{n-1}+i) & n > 1, \; i \le 0.
\end{cases}$$
%The definition of the fully discrete objective functions~\eqref{eq:obj_lms} as outlined in Section~\ref{sec:direct}, can be leveraged for this purpose. For simplicity, we only 
%outline the case for linear multistep schemes. 
\methodAcronym\ with an \spaceTimeAcronym\ trial subspace and the direct approach sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
\EP{I noticed you got rid of the assumption about the size of the generalized coordinates. I think we need to mention this, I put in a note below. Let me know if I'm missing something}
 \begin{equation}\label{eq:obj_gen_lms_final_st}
& \underset{\stgenstatey \in \RR{\stdimArg{n}}}{\text{minimize } }
\objectiveArgLMS{n}_{\text{D}} (\stbasisArg{n}{\timeWindowArg{n}{1}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n} ;\\
& \hspace{2. in}  \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots, 
 \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}}).%  \\
%where the space--time objective function for the direct method is defined as
%$$
%\objectiveArgLMS{n}_{\text{D-ST}}  :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}_{\text{D}}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots, 
% \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}}).
%$$
The boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace. 
Assuming $\text{Rank}(\stweightingMatOneArg{n}) \nstepsArg{n} \ge \stdimArg{n}$, the minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.
\begin{remark}
\methodAcronym\ with the direct solution approach and an \spaceTimeAcronym\ subspace comprises minimizing the same objective function as \methodAcronym\ with the direct solution approach and an \spatialAcronym\ trial subspace. 
\end{remark}
\KTC{I would add a remark that highlights the similarities between the direct
approach for the two kinds of subspaces. They minimize the exact same
objective function, which is interesting.} \EP{See if you agree with what I have}
\begin{remark}
For the limiting case where one window comprises the entire domain (i.e.,
	$\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial
	space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrix $\lspgWeightingST = \begin{bmatrix} \stweightingMatOneArg{1} & \cdots & \stweightingMatOneArg{1}\end{bmatrix}$
	\KTC{Don't we just need simply
	$W_{ST}=\text{diag}(W^1)$ (with $W^1$ the half-factor of $A^1$)?. I think this
	would be more general.}\EP{Yes, this is better. Not sure if diag is right though, see if you agree.}, and $\nstepsArg{1} = N_t$ time instances are employed that
	satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then
	$\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\
	with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG. 
\end{remark}

@@

\begin{remark}
\methodAcronym\ with \spaceTimeAcronym\ trial subspaces solved via the indirect approach naturally achieves ``collocation" in time as the full-order model residual needs to be queried at only the quadrature points. 
\end{remark}
\begin{remark}
For the limiting case where one window comprises the entire domain (i.e.,
	$\DeltaSlabArg{1}\equiv T$), the trial
	space is set to comprise the span of full solution trajectories, $\stspaceSTArg{1} = \text{span}\{ \stateFOMArgnt{}_i\}_{i=1}^{\stdimArg{1}}$ (where $\stateFOMArgnt{}_i$ are obtained, e.g.,  from training simulations at different parameter instances), and the weighting matrix is set to $\stweightingMatArg{1} = \mathbf{I}$, \methodAcronym\ with \spaceTimeAcronym\ trial subspaces and the indirect approach closely resembles the model reduction procedure proposed in~\cite{constantine_strom}; the approaches differ only in that Ref.~\cite{constantine_strom} places the constraint $\sum_{i=1}^{\stdimArg{1}} \stgenstateArg{1}_i = 1$. 
\end{remark}

\KTC{It would be great to a comment about how this is equivalent to Qiqi Wang
and Paul Constantine's method in the case of one time window and $\Pi^1$
corresponding to full solution trajectories. They use collocation-based
hyper-reduction at a few points in time (but for the full ODE residual at that
time), which you could mention. This is a very cool connection I didn't
realize before, and I think it's definitely worth mentioning!}
\EP{See if you agree with what I have. I worked with the trial space instead of the basis because we prescribe certain structure to the basis (such as the affine transformation)}.
%$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
%\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg]^T 
%\stweightingMatArg{n} 
%\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg].
%$$
%The stationary points are given by,
%\begin{multline*}\label{eq:st_stationary_discrete}
%\frac{\partial \objectiveDiscreteSTArg{n}}{\partial \stgenstatey}(\stgenstateArg{n}) = 
%\sum_{i=0}^{\ncollocSTArg{n}} \gamma_i \bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}}^T  - \stbasisArg{n}{\collocPointSTArg{n}{i}}^T \bigg[\frac{\partial
%\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n} +                    
%\stateInterceptArg{n})\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n} + \stateInterceptArg{n}) \bigg),\end{multline*}
%%subject to the boundary conditions 
%\begin{equation}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\approxstateIC &
%n=1. \end{cases} \end{equation} 
\subsection{\spaceTimeAcronym\ trial subspaces: summary}
\spaceTimeAcronym\ trial subspaces yield a series of space--time systems of algebraic equations over each window. As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial subspaces, 
a detailed exposition of solution techniques for these systems is not pursued here. It is sufficient to say that \methodAcronym\ with \spaceTimeAcronym\ trial subspaces yields a series of dense systems to be solved over each window.

>>>>

\subsection{\spaceTimeAcronym\ trial subspaces: direct solution approach}
The direct solution technique seeks to minimize the fully discrete objective
function associated with the FOM O$\Delta$E. We consider linear
multistep methods and leverage the discretization introduced in
Section~\ref{sec:direct}. For notational simplicity, we define an
index-mapping function that is equivalent to the mapping function
$\indexMapper$, but outputs only the first argument: 
$$\indexMappern: (n,i) \mapsto 
\begin{cases}
n & n = 1, \; i = 0, \\
n & n \ge 1, \; i > 0, \\
\indexMappern(n-1,\nstepsArg{n-1}+i) & n > 1, \; i \le 0.
\end{cases}$$
%The definition of the fully discrete objective functions~\eqref{eq:obj_lms} as outlined in Section~\ref{sec:direct}, can be leveraged for this purpose. For simplicity, we only 
%outline the case for linear multistep schemes. 
\methodAcronym\ with an \spaceTimeAcronym\ trial subspace and the direct approach sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
 \begin{equation}\label{eq:obj_gen_lms_final_st}
\begin{split}
& \underset{\stgenstatey \in \RR{\stdimArg{n}}}{\text{minimize } }
\objectiveArgLMS{n}_{\text{D}} (\stbasisArg{n}{\timeWindowArg{n}{1}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n} ;\\
& \hspace{2. in}  \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots, 
 \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}}).%  \\
%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
%n=1. \end{cases}
\end{split} 
\end{equation}
%where the space--time objective function for the direct method is defined as
%$$
%\objectiveArgLMS{n}_{\text{D-ST}}  :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}_{\text{D}}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots, 
% \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}}).
%$$
The boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace. 
Assuming $\text{Rank}(\stweightingMatOneArg{n}) \nstepsArg{n} \ge \stdimArg{n}$, the minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.
\begin{remark}
	Comparing optimization problems \eqref{eq:obj_gen_lms_final} and
	\eqref{eq:obj_gen_lms_final_st} reveals that
\methodAcronym\ with the direct solution approach minimizes the same objective
	function in the case of both \spaceTimeAcronym\ and \spatialAcronym\ trial
	subspaces.
\end{remark}
\begin{remark}
For the limiting case where one window comprises the entire domain (i.e.,
	$\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial
	subspace is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrix
	$\lspgWeightingST = \mathrm{diag}(\stweightingMatOneArg{1})$, and $\nstepsArg{1} = N_t$ time instances are employed that
	satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then
	$\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\
	with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG. 
\end{remark}

@@

\begin{remark}
\methodAcronym\ with \spaceTimeAcronym\ trial subspaces solved via the indirect approach naturally achieves ``collocation" in time as the full-order model residual needs to be queried at only the quadrature points. 
\end{remark}
\begin{remark}
For the limiting case where one window comprises the entire domain (i.e.,
	$\DeltaSlabArg{1}\equiv T$), the trial
	subspace is set to the span of full solution trajectories, $\stspaceSTArg{1}
	= \text{span}\{ \stateFOMArgnt{}_i\}_{i=1}^{\stdimArg{1}}$ (where
	$\stateFOMArgnt{}_i$ are obtained, e.g.,  from training simulations at
	different parameter instances), and the weighting matrix is set to
	$\stweightingMatArg{1} = \mathbf{I}$, \methodAcronym\ with
	\spaceTimeAcronym\ trial subspaces and the indirect approach closely
	resembles the model reduction procedure proposed in
	Ref.~\cite{constantine_strom}; the approaches differ only in that
	Ref.~\cite{constantine_strom} imposes the constraint
	$\sum_{i=1}^{\stdimArg{1}} \stgenstateArg{1}_i = 1$ in the associated
	minimization problem. 
\end{remark}
%$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
%\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg]^T 
%\stweightingMatArg{n} 
%\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg].
%$$
%The stationary points are given by,
%\begin{multline*}\label{eq:st_stationary_discrete}
%\frac{\partial \objectiveDiscreteSTArg{n}}{\partial \stgenstatey}(\stgenstateArg{n}) = 
%\sum_{i=0}^{\ncollocSTArg{n}} \gamma_i \bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}}^T  - \stbasisArg{n}{\collocPointSTArg{n}{i}}^T \bigg[\frac{\partial
%\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n} +                    
%\stateInterceptArg{n})\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n} + \stateInterceptArg{n}) \bigg),\end{multline*}
%%subject to the boundary conditions 
%\begin{equation}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\approxstateIC &
%n=1. \end{cases} \end{equation} 
\subsection{\spaceTimeAcronym\ trial subspaces: summary}
\spaceTimeAcronym\ trial subspaces yield a series of space--time systems of algebraic equations over each window. As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial subspaces, 
a detailed exposition of solution techniques for these systems is not pursued here. It is sufficient to say that \methodAcronym\ with \spaceTimeAcronym\ trial subspaces yields a series of dense systems to be solved over each window.
